Does an earthquake “know” how big it will be? A neural-net aided study

### 2024-06-02 article.tex
Replaced heavy figures with lighter versions.

MAGNET is composed of two main components. First, the catalog up to time $t$ (not including) is encoded by a long-short term memory (LSTM) unit, which produces a latent representation of the seismic history \yohai{cite LSTM paper}.
Second, this representation is combined with a space-time query specifying the coordinates, $\bf{x}_i$, $t_i$, of a future earthquake and passed into a fully connected neural network (FCNN) that produces a parametrized PDF of the event's magnitude, $p_{\textbf{x}_i, t_i}(m)$, see Fig. \ref{fig:intro_fig}a.

@@

The model parameters are optimized to maximize the log likelihood (LL) of the observed magnitudes. Importantly, during training we only supply the model with queries about spacetime coordinates where earthquakes indeed occurred, and only events above the completeness magnitude of the catalog are used as queries. A detailed description of the model's architecture is given in the methods section \neri{TODO}.

@@

\section{Results} \label{sec:results}
As explained above, the model outputs a PDF $p_{\textbf{x}_i, t_i}$ for each earthquake in the training set. We denote $\ell_i=\log\left(p_{\textbf{x}_i, t_i}(m_i)\right)$, the log-likelihood of the true magnitude $m_i$ as predicted by MAGNET.
The model is trained to maximize the average log-likelihood of the true magnitudes in the training set, $\mathcal{L} = -\langle \log{ \ell_i } \rangle$, where $\langle \cdot\rangle$ stands for the empirical average over all earthquakes in the data set.

% Focusing on existing events within the catalog due to our model's requirement for labeled data, our analysis utilizes the space-time coordinates of earthquakes as input. This necessitates excluding queries about arbitrary spatiotemporal points during training and evaluation, as these processes require knowledge of the true magnitude for each event at its specific location and time. Consequently, the raw output for our analysis consists of the model-generated PDFs of possible magnitudes for each event's time and location.

Figure \ref{fig:model_output}a shows the predicted PDFs for 100 randomly sampled events from the Southern California test set. The PDFs exhibit a clear trend: those that were calculated for higher-magnitude events are skewed towards larger magnitudes (warmer colors) compared to lower-magnitude events (cooler colors). This aligns with the expected behavior for an earthquake magnitude predictor. For comparison, the stationary GR distribution, the naive predictor which is independent of (recent) seismic history, is presented along the model's results.

This is a qualitative demonstration that the model works as expected. For a quantitative comparison, we calculate the $\mathcal{L}=-\langle \ell_i \rangle$ over all events in the test set, and benchmark it against state-of-the-art models:
the stationary Gutenberg-Richter (GR) distribution (fitted over the training set), and a few variants of a moving window GR approach~\cite{gulia_real-time_2019}. It is seen that MAGNET significantly and consistently outperforms all benchmarks across all three regions, as shown in Fig. \ref{fig:metrics}a-c. Further comparison to other benchmarks is given in the supplementary material \yohai{TODO: add the SM}.

In addition to comparing the log-likelihood of the observed data, one can convert the prediction problem to a binary question: will the earthquake at the spacetime query $\textbf{x}_i, t_i$ have be higher than a threshold $m_t$? This can be easily calculated from the prediction, as the probability that $m_i>m_t$ is simply $\int_{m_t}^{\infty}p(m')dm'$.
This task is more interpretable (will there be a large earthquake?) and can be evaluated using standard binary classification metrics such as area under receiver operator (ROC) curve \cite{Murphy} and under the interpolated precision-recall (PR) curve \cite{buttcher_information_2010} which suit better an imbalanced data set such as ours. Fig.~\ref{fig:metrics}d-f and g-i show these metrics for all three regions in examined and compare against the GR-variant models. Again, MAGNET outperforms all benchmarks.

To further assess performance in terms of predicted magnitudes, Figure \ref{fig:model_output}c presents the marginal probability density function (PDF) of magnitudes for MAGNET (red) compared to common benchmarks on the Southern California dataset. This figure reveals the general trends predicted by each model and facilitates comparison with the commonly used, stationary GR distribution. Furthermore, Fig. \ref{fig:model_output}c includes histograms of the train and test set magnitudes alongside the marginal PDFs.  We observe that the GR distribution closely follows the train set distribution, as expected for a stationary model.  In contrast, MAGNET aligns well with the test set distribution, suggesting its ability to generalize beyond the training data. Notably, dynamic GR benchmarks, exemplified by the "last 300 events" approach,  while exhibiting a better fit to the test set than the stationary GR, lack the detailed features captured by the model's PDF.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.5\textwidth]{figures/raw_results.pdf}
    \caption{
        \textbf{Benchmarking MAGNET}. \textbf{a}, PDFs produced by MAGNET for 100 randomly sampled events from the test set with train set's Gutenberg-Richter distribution (dashed black line). The true magnitude label, i.e. the true magntude of the event in query, is indicated by the color, interpreted by the colorbar at top. \textbf{b}, Likelihoods for returning the magnitude label by MAGNET (red), the 300 events moving window (orange) and the stationary GR (blue dashed line). \textbf{c}, Marginal magnitude distribution, i.e. stationary $p(m)$, produced by MAGNET and common benchmarks for the entire test set. Histograms of the train (and test) set are presented in orange (blue). \neri{c: Is there a more convenient way of showing the two superimposed distributions?}
    }
    \label{fig:model_output}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{figures/combined_barplots.pdf}
    \caption{
        \textbf{Metric scores for the MAGNET model, and common benchmark models. a}, \textbf{b}, \textbf{c}, Minus mean information content of our MAGNET model (red) and other common benchmark magnitude predictors (see labels on bars in the figure). A lower score indicated a better preforming model. \textbf{d, e, f}, Receiver Operating Characteristic (ROC) and \textbf{g, h, i} the interpolated precision-recall (PR) curve for a binary classifier determining the existence of a large ($m>=4$) event. The performance of such a classifier can be quantified by the area under the curve (AUC), noted in each frame, color coded identically to the bar plots. For the AUC metrics, a higher score indicated a better preforming classifier.
        }
        \label{fig:metrics}
\end{figure}

\begin{figure}[h!]
    \centering
    % \includegraphics[width=1\textwidth]{figures/info_over_time_hauksson.pdf}
    \caption{
        \textbf{Analysis of information gain of MAGNET over the benchmark. a,} Information gain of individual events in the Southern California test set. Scattered dots indicate magnitude at event index, colored by the information gain per event (colorbar indicated scale), here warm is an advantage for MAGNET, while cool is an advantage for the benchmark. Secondary horizontal axis (grey) indicates corresponds origin time. Cumulative Information gain (dashed black) and cumulative conditioned information gain (dashed grey) curves are superimposed on the scatter, demonstrating a constantly increasing function. \textbf{b-d}, Information gain (dashed black) and cumulative conditioned information gain (dashed grey) curves for Southern California, New Zealand and Japan test sets, respectively, all three regions examined in this study.\textbf{e-g}, Smoothed instantaneous information gain. Grey cloud shows the standard deviation of the smoothing window. All three curves are predominantly above the zero information gain line (dashed red). \textbf{h-i} Spatial distribution of information gain for all regions in study. Maps show the difference in mean information content per bin, model's advantage is indicated by warm colors, disadvantage by cool colors.
    }
    \label{fig:info_gain_over_time}
\end{figure}

The information gain achieved by our model for each event can be quantified by the difference in log-likelihood (LL) between our model's prediction and the baseline GR predictor:
\begin{equation}
    \Delta LL_i = \log{p_{\pmb{x_i}, t_i}^{(MAGNET)}(m_i)} - \log{p^{(GR)}(m_i)}
    \label{eq:information_gain}
\end{equation}
This equation measures the amount of information gained by using our MAGNET model rather than the GR benchmark, for a specific event. Analyzing this quantity per event across the entire test set reveals patterns indicating where and when our model achieves its advantage. Fig. \ref{fig:info_gain_over_time}a shows the test set for Southern California events magnitudes by their running index, colored according to the information gain as defined in Eq. \ref{eq:information_gain}. This time-domain representation allows us to track the cumulative information gain, calculated over the entire test set and displayed as the black dashed line in the figure. The predominantly increasing trend in cumulative information gain suggests that our model is consistently advantageous as opposed to a concentrated advantage in specific periods. Notably, for the Southern California data set, the most rapid information gain occurs during the aftershocks of the two Ridgecrest earthquakes. Surprisingly, even the pre-shock and main shock events demonstrate positive information gains over the GR benchmark, with values of $0.5$ and $1.4, \ \Delta LL$, respectively, although these events are typically considered completely unpredictable. This observation leads to a significant corollary: at least some portion of the information gain likely originates from a statistical signal embedded in the event sequence prior to event's occurrence. This point will be further elaborated upon in the discussion section.
\newline

@@

While the information gain observed in preshock and mainshocks presented above is encouraging, we address a potential concern regarding information gain during high temporal incompleteness periods\cite{stockman_forecasting_2023}. In order to account for the temporal incompleteness artifact, and factor it out of our information gain calculation, we recalculate the likelihood score of each event, $p(m_i)$ , conditioned on the magnitude exceeding a threshold, $\Tilde{m}$:

@@

For our analysis, we set $\Tilde{m} = m_c(t)$, determined dynamically using the maximum curvature method (Weimer and Weiss, 2000) within a window of 150 past and 150 future events. The resulting temporal incompleteness curves are presented in the SM \neri{TODO: add the $m_c(t)$ figure to SM}. After applying the renormalization of Eq. \ref{eq:conditioned_likelihood} the cumulative information gain is recalculated. The result is presented as the grey dashed curved in Fig.  \ref{fig:info_gain_over_time}a, as well as in Fig. \ref{fig:info_gain_over_time}b-d for all three tested regions. Evidently, all conditioned cumulative information gain curves show a steadily increasing trend, supporting the hypothesis that there is information content about the upcoming magnitude in the statistical properties of the catalog, excluding the temporal incompleteness artifact. The mean LL scores of the test set for each data catalog are presented in the supplementary material. \neri{TODO: add conditioned bar plots to SM}.

Beyond the temporal domain, MAGNET also exhibits an advantageous information gain distribution spatially (Fig. \ref{fig:info_gain_over_time}h-j). Each spatial bin's color represents the average log-likelihood difference between MAGNET's predictions and the common GR benchmark (as defined in Eq. \ref{eq:information_gain}, averaged and applied per bin). Warm colors indicate regions where MAGNET performs better. While the advantage in Southern California appears spatially diffuse, both New Zealand and Japan display concentrations of high advantage at the dataset edges, further from the mainland. We hypothesize this is linked to the high local magnitude of completeness, $m_c(x,y)$, in these regions. To address potential artifacts from spatially varying $m_c(x,y)$, we conducted further analysis. We computed local $m_c(x,y)$ per bin and selected only bins with completeness lower than the entire training set's completeness magnitude, $m_c \left( x,y \right)  \leq m_c^{ \left( train \right) }$. We then used these bins to create a minimal dataset and re-ran the model with all previous metrics. As in prior experiments, MAGNET outperformed benchmarks on this minimal dataset. Details and results of this analysis are provided in the supplementary material. \neri{TODO: add these results of minimal data set to the SM}.

@@

\newpage
\section*{Extended Data}
\begin{enumerate}
    \item Table: scores of all metrics - \textcolor{red}{\textbf{should I add this??}}
    \item Detailed architecture? - \textcolor{red}{\textbf{Make a nice version}}
    \item Colorful PDF examples for NZ and Japan - \textcolor{blue}{Done}
    \item Marginal distributions for NZ and Japan - \textcolor{blue}{Done}
    \item Colored scatter magnitude vs index, NZ and Japan - \textcolor{blue}{\textbf{Done}}
    \item $m_c(t)$ for all three regions - \textcolor{blue}{Done}
    \item Bar plots of conditioned $<LL>$ - \textcolor{blue}{Done}
    \item Spatial $m_c(x, y)$ and low $m_c(x, y) \leq m_c^{train}$ bins - \textcolor{blue}{Done}
\end{enumerate}

@@

\newpage
\begin{table}[]
    \centering
    \includegraphics[width=1\textwidth]{figures/mean_ll_table.pdf}
    \caption{Mean log likelihood, $\mathcal{L}$, for additional tested benchmarks.}
    \label{tab:mean_ll_all_benchmarks}
\end{table}[h!]

\newpage
\begin{table}[]
    \centering
    \includegraphics[width=1\textwidth]{figures/mean_ll_cond_table.pdf}
    \caption{Mean log likelihood, $\mathcal{L}$, conditioned on temporal incompletness, for additional tested benchmarks.}
    \label{tab:mean_ll_all_benchmarks_conditioned}
\end{table}[h!]

@@

\newpage
\begin{figure}[h!]
	\centering
        \includegraphics[width=1\textwidth]{figures/info_gain_scatter_NZ_and_Japan.pdf}
	\caption{
            \textbf{Information gain over time domain for New Zealand and Japan data sets.} This figure presents results for the data sets not displayed in Fig. \ref{fig:info_gain_over_time}a. \textbf{a} Entire test set of New Zealand and a selected time span (\textbf{b}) focusing on the 2016 Te Araroa earthquake and the 2016 Kaikōura earthquake. \textbf{c} The entire Japan data set and a focused time span \textbf{d} of the 2011 Tōhoku earthquake and some of its aftershocks.
         }
\label{fig:nz_japan_info_gain}
\end{figure}

\newpage
\begin{figure}[h!]
    \centering
        % \includegraphics[width=1\textwidth]{figures/temporal_incompleteness.pdf}
    \caption{
    \textbf{Temporal Incompleteness for test sets of all regions examined.} Calculated temporal incompleteness for test sets is presented in red, superimposed over the events above (black) and below (grey) the completeness magnitude of the train set, $m_c^{(train)}$. Events are plotted by the index of their appearance in the test set. Events of low magnitude, $m_c^{(train)}$ (grey), that are not included in the evaluation, are plotted at their interpolated location along the horizontal axis. Data for Japan (c) is displayed only near the 2011 Tōhoku earthquake to reduce data density and enhance clarity of the figure.
    }
    \label{fig:temp_incompleteness}
\end{figure}

>>>>

MAGNET is composed of two main components. First, the catalog up to time $t$ (not including) is encoded by a long-short term memory (LSTM) unit, which produces a latent representation of the seismic history \neri{cite Hochreiter, S. and Schmidhuber, J. ürgen. Long short-term memory. Neural Comput. 9, 1735–1780 (1997).}.
Second, this representation is combined with a space-time query specifying the coordinates, $\bf{x}_i$, $t_i$, of a future earthquake and passed into a fully connected neural network (FCNN) that produces a parametrized PDF of the event's magnitude, $p_{\textbf{x}_i, t_i}(m)$, see Fig. \ref{fig:intro_fig}a.

@@

The model parameters are optimized to maximize the log likelihood (LL) of the observed magnitudes. Importantly, during training we only supply the model with queries about spacetime coordinates where earthquakes indeed occurred, and only events above the completeness magnitude of the catalog are used as queries. A detailed description of the model's architecture is given in the methods section.

@@

\section{Results} \label{sec:results}
As explained above, the model outputs a PDF $p_{\textbf{x}_i, t_i}(m_i)$ for each earthquake in the training set. We denote $\ell_i=\log\left(p_{\textbf{x}_i, t_i}(m_i)\right)$, the log-likelihood of the true magnitude $m_i$ as predicted by MAGNET.  The values of $\ell_i$ for MAGNET and the leading benchmark are presented in Extended Material Fig. \ref{fig:labels_likelihood}.

The model is trained to maximize the average log-likelihood of the true magnitudes in the training set, $\mathcal{L} = -\langle \ell_i \rangle$, where $\langle \cdot\rangle$ stands for the empirical average over all earthquakes in the data set. \neri{I think the two equations in this paragraph should be numbered.}

% Focusing on existing events within the catalog due to our model's requirement for labeled data, our analysis utilizes the space-time coordinates of earthquakes as input. This necessitates excluding queries about arbitrary spatiotemporal points during training and evaluation, as these processes require knowledge of the true magnitude for each event at its specific location and time. Consequently, the raw output for our analysis consists of the model-generated PDFs of possible magnitudes for each event's time and location. 
\neri{I liked this sentence: \textit{Here we demonstrate that MAGNET outperforms all benchmarks in predicting earthquake magnitude, demonstrating that magnitudes are not statistically independent on seismic history.} Don't you think it's good to start with this clam to focus the reader on the objective?}\yohai{yes I do, which is why this is the last sentence in the previous paragraph}
Figure \ref{fig:model_output}a shows the predicted PDFs for 100 randomly sampled events from the Southern California test set. The PDFs exhibit a clear trend: those that were calculated for higher-magnitude events are skewed towards larger magnitudes (warmer colors) compared to lower-magnitude events (cooler colors). This aligns with the expected behavior for an earthquake magnitude predictor. For comparison, the stationary GR distribution, the naive predictor which is independent of (recent) seismic history, is presented along the model's results.

This is a qualitative demonstration that the model works as expected. For a quantitative comparison, we calculate the $\mathcal{L}=-\langle \ell_i \rangle$ over all events in the test set, and benchmark it against state-of-the-art models:
the stationary Gutenberg-Richter (GR) distribution (fitted over the training set), and a few variants of a moving window GR approach~\cite{gulia_real-time_2019}. It is seen that MAGNET significantly and consistently outperforms all benchmarks across all three regions, as shown in Fig. \ref{fig:metrics}a-c. Further comparison to other benchmarks is given in Extended Data Table \ref{tab:mean_ll_all_benchmarks}.

In addition to comparing the log-likelihood of the observed data, one can convert the prediction problem to a binary question: will the earthquake at the spacetime query $\textbf{x}_i, t_i$ have be higher than a threshold $m_t$? This can be easily calculated from the prediction, as the probability that $m_i>m_t$ is simply $\int_{m_t}^{\infty}p(m')dm'$.
This task, asking whether a given event will be a large earthquake, is more interpretable and can be evaluated using standard binary classification metrics such as area under receiver operator (ROC) curve \cite{Murphy} and under the interpolated precision-recall (PR) curve \cite{buttcher_information_2010} which suit better an imbalanced data set such as ours. Fig.~\ref{fig:metrics}d-f and g-i show these metrics for all three regions in examined and compare against the GR-variant models. Again, MAGNET outperforms all benchmarks.

Finally, Figure~\ref{fig:model_output}b presents the marginal probability density function (PDF) of magnitudes for MAGNET's prediction, $P(m)=N^{-1}\sum_i p_{\textbf{x}_i, t_i}(m)$. This represents the average probability distribution of magnitudes over the whole test set. It is seen that although the model was not constrained to do so, the average distribution aligns well with the ``uninformed'' GR distribution. That is, although MAGNET's prediction for each individual earthquake may deviate significantly from the GR distribution, it does reproduce this well known long-time behavior on average. It is also seen that the average prediction of the 300-day moving window GR deviates significantly from thte GR distribution. We also observe that MAGNET's average prediction aligns well with the \emph{test set} GR distribution, rather than the training set, suggesting its ability to generalize beyond the training data. This observation is noticeable even more so in the New Zealand and Japan data sets presented in Extended Data Fig. \ref{fig:model_output_em}.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.5\textwidth]{figures/raw_results.pdf}
    \caption{
        \textbf{MAGNET's output}. \textbf{a}, PDFs produced by MAGNET for 100 randomly sampled events from the test set, train set's Gutenberg-Richter distribution superimposed (dashed black line). The true magnitude label, i.e. the true magntude of the event in query, is indicated by the color, interpreted by the colorbar at top. \textbf{b}, Marginal magnitude distribution, i.e. stationary $p(m)$, produced by MAGNET and common benchmarks for the entire test set. Histograms of the train (and test) set are presented in orange (blue).
    }
    \label{fig:model_output}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{figures/combined_barplots.pdf}
    \caption{
        \textbf{Benchmarking MAGNET. a}, \textbf{b}, \textbf{c}, Minus mean information content of our MAGNET model (red) and other common benchmark magnitude predictors (see labels on bars in the figure). A lower score indicates a better preforming model. \textbf{d, e, f}, Receiver Operating Characteristic (ROC) and \textbf{g, h, i} the interpolated precision-recall (PR) curve for a binary classifier determining the existence of a large ($m>=4$) event. The performance of such a classifier can be quantified by the area under the curve (AUC), noted in each frame, color coded identically to the bar plots. For the AUC metrics, a higher score indicates a better preforming classifier.
        }
        \label{fig:metrics}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{figures/info_over_time_hauksson.pdf}
    \caption{
        \textbf{Information gain in MAGNET. a,} Information gain of individual events in the Southern California test set. Scattered dots indicate magnitude at event index, colored by the information gain per event (scale indicated by colorbar), here warm is an advantage for MAGNET, while cool is an advantage for the benchmark. Secondary horizontal axis (grey) indicates corresponds origin time. Cumulative Information gain (dashed black) and cumulative conditioned information gain (dashed grey) curves are superimposed on the scatter, demonstrating a constantly increasing function. \textbf{b-d}, Information gain (dashed black) and cumulative conditioned information gain (dashed grey) curves for Southern California, New Zealand and Japan test sets, respectively, all three regions examined in this study.\textbf{e-g}, Smoothed instantaneous information gain. Grey cloud shows the standard deviation of the smoothing window. All three curves are predominantly above the zero information gain line (dashed red). \textbf{h-i} Spatial distribution of information gain for all regions in study. Maps show the difference in mean information content per bin, model's advantage is indicated by warm colors, disadvantage by cool colors.
    }
    \label{fig:info_gain_over_time}
\end{figure}

It is interesting to ask The information gain achieved by our model for each event can be quantified by the difference in log-likelihood (LL) between our model's prediction and the baseline GR predictor:
\begin{equation}
    \Delta LL_i = \log{p_{\pmb{x_i}, t_i}^{(MAGNET)}(m_i)} - \log{p^{(GR)}(m_i)}
    \label{eq:information_gain}
\end{equation}
This equation measures the amount of information gained by using our MAGNET model rather than the GR benchmark, for a specific event. Analyzing this quantity per event across the entire test set reveals patterns indicating where and when our model achieves its advantage. Fig. \ref{fig:info_gain_over_time}a shows the test set for Southern California events magnitudes by their running index, colored according to the information gain as defined in Eq. \ref{eq:information_gain}. This time-domain representation allows us to track the cumulative information gain, calculated over the entire test set and displayed as the black dashed line in the figure. The predominantly increasing trend in cumulative information gain suggests that our model is consistently advantageous as opposed to a concentrated advantage in specific periods. Notably, for the Southern California data set, the most rapid information gain occurs during the aftershocks of the two Ridgecrest earthquakes. Surprisingly, even the pre-shock and main shock events demonstrate positive information gains over the GR benchmark, with values of $0.5$ and $1.4, \ \Delta LL$, respectively, although these events are typically considered completely unpredictable. This observation leads to a significant corollary: at least some portion of the information gain likely originates from a statistical signal embedded in the event sequence prior to event's occurrence. This point will be further elaborated upon in the discussion section.
\newline

@@

\section{Assessment of Measurement Reliability}
\neri{YOHAI: I've added this title above. thoughts?}\newline
While the information gain observed in preshock and mainshocks presented above is encouraging, we address a potential concern regarding information gain during high temporal incompleteness periods\cite{stockman_forecasting_2023}. In order to account for the temporal incompleteness artifact, and factor it out of our information gain calculation, we recalculate the likelihood score of each event, $p(m_i)$ , conditioned on the magnitude exceeding a threshold, $\Tilde{m}$:

@@

For our analysis, we set $\Tilde{m} = m_c(t)$, determined dynamically using the maximum curvature method \cite{wiemer_minimum_2000} within a window of 150 past and 150 current and future events. The resulting temporal incompleteness curves are presented in the Extended Material Fig. \ref{fig:temp_incompleteness}. After applying the renormalization of Eq. \ref{eq:conditioned_likelihood} the cumulative information gain is recalculated. The result is presented as the grey dashed curved in Fig.  \ref{fig:info_gain_over_time}a, as well as in Fig. \ref{fig:info_gain_over_time}b-d for all three tested regions. Evidently, all conditioned cumulative information gain curves show a steadily increasing trend, supporting the hypothesis that there is information content about the upcoming magnitude in the statistical properties of the catalog, excluding the temporal incompleteness artifact. A bar plot comparison of the conditioned mean LL scores are presented in Extended Material Fig. \ref{fig:mean_ll_mc_conditioned}. Extended Material Table \ref{tab:mean_ll_all_benchmarks_conditioned} displays the conditioned scores for all benchmarks tested in this study. \neri{Perhaps the barplot is unecessary?}

Beyond the temporal domain, MAGNET also exhibits an advantageous information gain distribution spatially (Fig. \ref{fig:info_gain_over_time}h-j). Each spatial bin's color represents the average log-likelihood difference between MAGNET's predictions and the common GR benchmark (as defined in Eq. \ref{eq:information_gain}, averaged and applied per bin). Warm colors indicate regions where MAGNET performs better. While the advantage in Southern California appears spatially diffuse, both New Zealand and Japan display concentrations of high advantage at the dataset edges, further from the mainland. We hypothesize this is linked to the high local magnitude of completeness, $m_c(x,y)$, in these regions. To address potential artifacts from spatially varying $m_c(x,y)$, we conducted further analysis. We computed local $m_c(x,y)$ per bin and selected only bins with completeness lower than the entire training set's completeness magnitude, $m_c \left( x,y \right)  \leq m_c^{ \left( train \right) }$. We then used these bins to create a minimal dataset and tested the models performance on it. MAGNET holds its advantage in the Southern California and New Zealand datasets. However, MAGNET loses when this analysis is preformed on the Japan dataset. To investigate further, we choose a another subset of the Japan catalog, defined by a tight polygon encompassing the majority of bins that hold. $m_c \left( x,y \right)  \leq m_c^{ \left( train \right) }$ bins. This results in mostly the territory of Honshu, Kyushu and Shikoku islands. We then train a new model for this subset, which, as in the previous experiments, displays MAGNET's superiority for magnitude prediction. We conclude from this procedure that constructing a magnitude prediction model for regions with high variability of $m_c(x,y)$ \textit{is possible}, though it might consist of training different regions separately. Details and results of this analysis are provided in Extended Material Fig. \ref{fig:mean_ll_conditioned_spatial_mc}.

\section{Discussion}
\neri{YOHAI LET'S MEET TO WRITE CMON}

@@

\newpage
\section*{Extended Data}
\begin{enumerate}
    \item Table: scores of all metrics - \textcolor{blue}{Done}
    \item Detailed architecture? - \textcolor{red}{\textbf{Make a nice version}}
    \item Colorful PDF examples for NZ and Japan - \textcolor{blue}{Done}
    \item Marginal distributions for NZ and Japan - \textcolor{blue}{Done}
    \item Colored scatter magnitude vs index, NZ and Japan - \textcolor{blue}{\textbf{Done}}
    \item $m_c(t)$ for all three regions - \textcolor{blue}{Done}
    \item Bar plots of conditioned $<LL>$ - \textcolor{blue}{Done}
    \item Spatial $m_c(x, y)$ and low $m_c(x, y) \leq m_c^{train}$ bins - \textcolor{blue}{Done}
\end{enumerate}

@@

\newpage
\begin{table}[h!]
    \centering
    \includegraphics[width=1\textwidth]{figures/mean_ll_table.pdf}
    \caption{Mean log likelihood score, $\mathcal{L}$, for various tested benchmarks.}
    \label{tab:mean_ll_all_benchmarks}
\end{table}

\newpage
\begin{table}[h!]
    \centering
    \includegraphics[width=1\textwidth]{figures/mean_ll_cond_table.pdf}
    \caption{Mean log likelihood score, $\mathcal{L}$, as the result is conditioned to be above the temporal incompletness.}
    \label{tab:mean_ll_all_benchmarks_conditioned}
\end{table}

@@

\newpage
\begin{figure}[h!]
	\centering
        \includegraphics[width=1\textwidth]{figures/info_gain_scatter_NZ_and_Japan.pdf}
	\caption{
            \textbf{Information gain of MAGNET over the benchmark, New Zealand and Japan data sets.} Scattered dots indicate magnitude at event index, colored by the information gain per event. \textbf{a} Entire test set of New Zealand and a selected time span (\textbf{b}) focusing on the 2016 Te Araroa earthquake and the 2016 Kaikōura earthquake. \textbf{c} The entire Japan data set and a focused time span \textbf{d} of the 2011 Tōhoku earthquake and some of its aftershocks.
         }
\label{fig:nz_japan_info_gain}
\end{figure}

\newpage
\begin{figure}[h!]
    \centering
        \includegraphics[width=1\textwidth]{figures/temporal_incompleteness.pdf}
    \caption{
    \textbf{Temporal Incompleteness for test sets of all regions examined.} Calculated temporal incompleteness for test sets is presented in red, superimposed over the events above (black) and below (grey) the completeness magnitude of the train set, $m_c^{(train)}$. Events are plotted by the index of their appearance in the test set. Events of low magnitude, $m_c^{(train)}$ (grey), that are not included in the evaluation, are plotted at their interpolated location along the horizontal axis. Data for Japan (c) is displayed only near the 2011 Tōhoku earthquake to reduce data density and enhance clarity of the figure.
    }
    \label{fig:temp_incompleteness}
\end{figure}
